---
title: "knb-lter-cap.624"
author: "SRE"
date: "March 13, 2018"
output: html_document
---

This version 624.2 is a new workflow based on the Rmd format. There are new data available but, in the interest of time for the review, only data currently in the database at the time of construction are going into this version 2. More recent data and the discharge project stuff at the Salt River sites need to go into a version 3 soon.

Note that this new workflow if pretty-much hit the button and everything should be built. Still not addressed at in this current configuration, however, are the keyword attributes, so those still have to be done by hand. Note also that the kml file is going to be overwritten every time.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries}
library(EML)
library(RPostgreSQL)
library(RMySQL)
library(tidyverse)
library(tools)
library(readxl)
library(aws.s3)
library(capeml)
```

```{r dataset_details}
projectid <- 624
packageIdent <- 'knb-lter-cap.624.2'
pubDate <- '2018-03-14'
```
 
```{r helper_functions}
source('~/localRepos/reml-helper-tools/writeAttributesFn.R')
source('~/localRepos/reml-helper-tools/createDataTableFromFileFn.R')
source('~/localRepos/reml-helper-tools/createKMLFn.R')
source('~/localRepos/reml-helper-tools/address_publisher_contact_language_rights.R')
source('~/localRepos/reml-helper-tools/createOtherEntityFn.R')
source('~/localRepos/reml-helper-tools/createPeople.R')
source('~/localRepos/reml-helper-tools/createFactorsDataframe.R')
source('~/localRepos/reml-helper-tools/amazon_file_upload.R')
```

```{r connections::amazon}
source('~/Documents/localSettings/aws.s3')
```

```{r connections::postgres::local, eval=FALSE}
source('~/Documents/localSettings/pg_local.R')
pg <- pg_local
```

```{r connections::postgres::prod, eval=T }
source('~/Documents/localSettings/pg_prod.R')
pg <- pg_prod
```

```{r connections::mysql::prod, eval=T }
source('~/Documents/localSettings/mysql_prod.R')
prod <- mysql_prod
```

```{r runoff_chemistry}

# data from 2007 (single storm at IBW) omitted owing to lack of times 61 values
# sans an analysis id are omitted

runoffChemistry <- dbGetQuery(pg, "
  SELECT
    s.bottle,
    sites.abbreviation AS runoff_location,
    s.sample_datetime AS runoff_datetime,
    r.replicate,
    a.analysis_name AS analysis,
    r.concentration,
    dq.data_qualifier_label AS data_qualifier,
    r.comments
FROM
  stormwater.results r
  JOIN stormwater.samples s ON (s.sample_id = r.sample_id)
  JOIN stormwater.analysis a ON (r.analysis_id = a.analysis_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier)
WHERE
  EXTRACT (YEAR FROM s.sample_datetime) > 2007 AND
  r.analysis_id IS NOT NULL AND
  (bottle NOT ILIKE '%rain%' OR bottle IS NULL);")

# need to add a definitive reference for whether a sample is a blank given that
# bottle is omitted (and confusing)
runoff_chemistry <- runoffChemistry %>%
  mutate(
    blank = as.factor(ifelse(grepl('blk|blank', bottle, ignore.case = T), 'TRUE', 'FALSE')),
    runoff_location = as.factor(runoff_location),
    analysis = as.factor(analysis)
  ) %>%
  select(-bottle)

# writeAttributes(runoff_chemistry) # write data frame attributes to a csv in current dir to edit metadata

runoff_chemistry_desc <- "stormwater runoff chemistry during runoff-generating storms at CAP LTER stormwater sampling sites"

# runoff_chemistry_factors <- factorsToFrame(runoff_chemistry)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
runoff_chemistry_DT <- createDTFF(dfname = runoff_chemistry,
                                  # factors = runoff_chemistry_factors,
                                  description = runoff_chemistry_desc,
                                  dateRangeField = 'runoff_datetime')

```

```{r rainfall_chemistry}

rainfallChemistry <- dbGetQuery(pg, "
SELECT
  sites.abbreviation AS rain_location,
  s.sample_datetime AS rain_datetime,
  r.replicate,
  a.analysis_name AS analysis,
  r.concentration,
  dq.data_qualifier_label AS data_qualifier,
  r.comments
FROM
  stormwater.results r
  JOIN stormwater.samples s ON (s.sample_id = r.sample_id)
  JOIN stormwater.analysis a ON (r.analysis_id = a.analysis_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier)
WHERE
  EXTRACT (YEAR FROM s.sample_datetime) > 2007 AND
  r.analysis_id IS NOT NULL AND
  bottle ILIKE '%rain%'
ORDER BY sites.abbreviation, s.sample_datetime, a.analysis_name;")

rainfall_chemistry <- rainfallChemistry %>%
  mutate(
    rain_location = as.factor(rain_location),
    analysis = as.factor(analysis),
    rain_datetime = format(rain_datetime, format = "%Y-%m-%d")
  )

# writeAttributes(rainfall_chemistry) # write data frame attributes to a csv in current dir to edit metadata

rainfall_chemistry_desc <- "water chemistry of collected rainfall at CAP LTER stormwater sampling sites"

# rainfall_chemistry_factors <- factorsToFrame(rainfall_chemistry)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
rainfall_chemistry_DT <- createDTFF(dfname = rainfall_chemistry,
                                    # factors = rainfall_chemistry_factors,
                                    description = rainfall_chemistry_desc,
                                    dateRangeField = 'rain_datetime')

```

```{r rainfall}

rainfall <- dbGetQuery(pg,"
SELECT
  sites.abbreviation AS rain_location,
  r.event_datetime AS rain_datetime,
  r.rainfall_quantity AS rain_quantity
FROM
  stormwater.rainfall r
  JOIN stormwater.sites ON (r.site_id = sites.site_id);")

rainfall <- rainfall %>%
  mutate(
    rain_location = as.factor(rain_location)
  )

# writeAttributes(rainfall) # write data frame attributes to a csv in current dir to edit metadata

rainfall_desc <- "depth of precipitation as measured by tipping bucket rain gauge at CAP LTER stormwater sampling sites"

# rainfall_factors <- factorsToFrame(rainfall)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
rainfall_DT <- createDTFF(dfname = rainfall,
                          # factors = rainfall_factors,
                          description = rainfall_desc,
                          dateRangeField = 'rain_datetime')
```

```{r discharge}

discharge <- dbGetQuery(pg, "
SELECT
  sites.abbreviation AS discharge_location,
  r.event_datetime AS discharge_datetime,
  r.water_height AS water_height,
  r.discharge AS raw_discharge,
  r.discharge_corrected AS edited_discharge
FROM
  stormwater.discharge r
  JOIN stormwater.sites ON (r.site_id = sites.site_id)
;")

discharge <- discharge %>%
  mutate(
    discharge_location = as.factor(discharge_location)
  )

# writeAttributes(discharge) # write data frame attributes to a csv in current dir to edit metadata

discharge_desc <- "discharge at CAP LTER stormwater sampling sites"

# discharge_factors <- factorsToFrame(discharge)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
discharge_DT <- createDTFF(dfname = discharge,
                           # factors = discharge_factors,
                           description = discharge_desc,
                           dateRangeField = 'discharge_datetime')
```

```{r particulates}

particulates <- dbGetQuery(pg, "
SELECT
  s.bottle,
  sites.abbreviation AS runoff_location,
  s.sample_datetime AS runoff_datetime,
  r.replicate AS replicate,
  r.filter_initial AS filter_wt,
  r.filter_dry AS filter_dry_wt,
  r.volume_filtered AS vol_filtered,
  r.filter_ashed AS filter_ash_wt,
  r.data_qualifier AS data_qualifier,
  r.comments AS comments
FROM
  stormwater.solids r
  JOIN stormwater.samples s ON (s.sample_id = r.sample_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier);")

particulates <- particulates %>%
  mutate(
    blank = as.factor(ifelse(grepl('blk|blank', bottle, ignore.case = T), 'TRUE', 'FALSE')),
    runoff_location = as.factor(runoff_location),
    data_qualifier = as.character(data_qualifier)
  ) %>%
  select(-bottle)

# writeAttributes(particulates) # write data frame attributes to a csv in current dir to edit metadata

particulates_desc <- "mass of particulates in stormwater during runoff-generating storms at CAP LTER stormwater sampling sites"

# particulates_factors <- factorsToFrame(particulates)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
particulates_DT <- createDTFF(dfname = particulates,
                              # factors = particulates_factors,
                              description = particulates_desc,
                              dateRangeField = 'runoff_datetime')

```

```{r sample_locations}

sampling_location <- dbGetQuery(pg,"
SELECT
  abbreviation AS runoff_location,
  site_name AS name,
  latitude,
  longitude
FROM stormwater.sites;")

sampling_location <- sampling_location %>%
  mutate(
    runoff_location = as.factor(runoff_location)
  )

# writeAttributes(sampling_location) # write data frame attributes to a csv in current dir to edit metadata

sampling_location_desc <- "catalog of CAP LTER stormwater sampling locations"

# sampling_location_factors <- factorsToFrame(sampling_location)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
sampling_location_DT <- createDTFF(dfname = sampling_location,
                                   # factors = sampling_location_factors,
                                   description = sampling_location_desc)
```

```{r analyses}

analytes <- dbGetQuery(pg,"
SELECT
  DISTINCT ON (analysis_name)
  analysis_name as analysis,
  analysis_description AS description,
  units,
  instrument
FROM
  stormwater.analysis
  RIGHT JOIN stormwater.results r ON (r.analysis_id = analysis.analysis_id)
WHERE analysis_name IS NOT NULL;")


# writeAttributes(analytes) # write data frame attributes to a csv in current dir to edit metadata

analytes_desc <- "catalog and details of water chemistry analytes measured as part of CAP LTER stormwater monitoring"

# analytes_factors <- factorsToFrame(analytes)

# create data table based on metadata provided in the companion csv
# use createdataTableFn() if attributes and classes are to be passed directly
analytes_DT <- createDTFF(dfname = analytes,
                          # factors = analyses_factors,
                          description = analytes_desc)
```

```{r catchments}

stormwater_catchments <- createKML(kmlobject = 'stormwater_catchments.kml',
                                   description = "The location and configuration of watersheds that are or have been sampled as part of the CAP LTER's research on stormwater monitoring in the greater Phoenix metropolitan area. Catchments for sampling locations along the Salt River (centralNorth, centralSouth, Ave7th) and Price have not been delineated by the CAP LTER.")

```

```{r title}
title <- 'Long-term monitoring of stormwater runoff and water quality in urbanized watersheds of the greater Phoenix metropolitan area, ongoing since 2008'
```

```{r abstract}

# abstract from file or directly as text
abstract <- as(set_TextType("abstract.md"), "abstract") 
# abstract <- 'abstract text'
```

```{r people}

# creators
danChilders <- addCreator('d', 'childers')
stevanEarl <- addCreator('s', 'earl')
nancyGrimm <- addCreator('n', 'grimm')
rebeccaHale <- addCreator('r', 'hale')
lauraTurnbull <- addCreator('l', 'turnbull')

creators <- c(as(danChilders, 'creator'),
              as(stevanEarl, 'creator'),
              as(nancyGrimm, 'creator'),
              as(rebeccaHale, 'creator'),
              as(lauraTurnbull, 'creator'))

# metadata providers
stevanEarl <- addMetadataProvider('s', 'earl')
quincyStewart <- addMetadataProvider('q', 'stewart')
sallyWittlinger <- addMetadataProvider('s', 'wittlinger')

metadataProvider <-c(as(stevanEarl, 'metadataProvider'),
                     as(quincyStewart, 'metadataProvider'),
                     as(sallyWittlinger, 'metadataProvider'))

```

```{r keywords}

# CAP IRTs for reference: https://sustainability.asu.edu/caplter/research/
# be sure to include these as appropriate

keywordSet <-
  c(new("keywordSet",
        keywordThesaurus = "LTER controlled vocabulary",
        keyword =  c("nutrients",
                     "dissolved organic carbon",
                     "dissolved organic nitrogen",
                     "dissolved inorganic nitrogen",
                     "rivers",
                     "discharge",
                     "water chemistry",
                     "particulates",
                     "nitrate",
                     "nitrite",
                     "ammonium",
                     "anions",
                     "cations",
                     "chloride",
                     "phosphorus",
                     "nitrogen",
                     "specific conductivity",
                     "urban",
                     "runoff",
                     "watersheds",
                     "total dissolved nitrogen")),
    new("keywordSet",
        keywordThesaurus = "LTER core areas",
        keyword =  c("water and fluxes",
                     "parks and rivers",
                     "movement of organic matter",
                     "movement of inorganic matter")),
    new("keywordSet",
        keywordThesaurus = "Creator Defined Keyword Set",
        keyword =  c("stormwater",
                     "catchment")),
    new("keywordSet",
        keywordThesaurus = "CAPLTER Keyword Set List",
        keyword =  c("cap lter",
                     "cap",
                     "caplter",
                     "central arizona phoenix long term ecological research",
                     "arizona",
                     "az",
                     "indian bend wash",
                     "salt river",
                     "arid land"))
  )

```

```{r methods}

methods <- set_methods("methods.md")
```

```{r coverages}

begindate <- format(min(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
enddate <- format(max(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
geographicDescription <- "CAP LTER study area"
coverage <- set_coverage(begin = begindate,
                         end = enddate,
                         geographicDescription = geographicDescription,
                         west = -112.082, east = -111.865,
                         north = +33.6149, south = +33.3745)
```

```{r construct_dataset}

# from sourced file:
  # address
  # publisher
  # contact
  # rights
  # distribution

# generate a list of EML dataTables
listOfDataTables <- lapply(ls(pattern = "_DT"), function(DT) { get(DT) } )

# print list as a safety step
print(ls(pattern= "_DT"))

# DATASET
dataset <- new("dataset",
               title = title,
               creator = creators,
               pubDate = pubDate,
               metadataProvider = metadataProvider,
               # associatedParty = associatedParty,
               intellectualRights = rights,
               abstract = abstract,
               keywordSet = keywordSet,
               coverage = coverage,
               contact = contact,
               methods = methods,
               distribution = metadata_dist,
               dataTable = listOfDataTables,
               otherEntity = c(stormwater_catchments)) # if other entity is relevant

               # dataTable = c(first_DT,
               #               second_DT))

```

```{r custom_units, eval=FALSE}

```

```{r construct_eml}

if(exists('unitList')) {
  eml <- new("eml",
             schemaLocation = "eml://ecoinformatics.org/eml-2.1.1  http://nis.lternet.edu/schemas/EML/eml-2.1.1/eml.xsd",
             packageId = packageIdent,
             scope = "system",
             system = "knb",
             access = lter_access,
             dataset = dataset,
             additionalMetadata = as(unitList, "additionalMetadata"))
} else {
  eml <- new("eml",
             schemaLocation = "eml://ecoinformatics.org/eml-2.1.1  http://nis.lternet.edu/schemas/EML/eml-2.1.1/eml.xsd",
             packageId = packageIdent,
             scope = "system",
             system = "knb",
             access = lter_access,
             dataset = dataset)
}

```

```{r write_eml}

# write the eml to file
write_eml(eml, paste0(packageIdent, ".xml"))
```

```{r preview_data_file_to_upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(projectid, "_"))
```

```{r upload_data_S3}

# upload files to S3
lapply(list.files(pattern = paste0(projectid, "_")), dataToAmz)
```

```{r clean_up}

# remove data files
dataFilesToRemove <- dir(pattern = paste0(projectid, "_"))
file.remove(dataFilesToRemove)

# EML to S3
if(length(list.files(pattern = "*.xml")) == 1) {
  emlToAmz(list.files(pattern = "*.xml")) } else {
    print("more than one xml file found")
  }

# EML to cap-data/cap-data-eml and remove file from project
if(length(list.files(pattern = "*.xml")) == 1) {
  file.copy(list.files(pattern = "*.xml"), "/home/srearl/localRepos/cap-data/cap-data-eml")
  file.remove(list.files(pattern = "*.xml")) } else {
    print("more than one xml file found")
  }
```

